---
title: "Other patients"
author: "Zach Wang"
date: "9/8/2020"
output:
  html_document: default
  word_document: default
  pdf_document: default
---
## 1. Function Implementations
### 1.1 function 1: *ReadFile* and attach packages
```{r setup, include=FALSE}
library(zoo)
library(fda.usc)
library(dplyr)
library(reshape)
library(reshape2)
library(pracma)
library(ggplot2)

## function 1:
# function to read data from specified file location, with respect to Node and Time of interest
ReadFile <- function(file_loc, time_subset, node_subset){
  data_mat <- read.csv(file_loc, header=TRUE)
  data_mat <- select(data_mat, -c(X)) # input data åhas an extra column "X"
  return(as.matrix(data_mat[time_subset,node_subset]))
}
```

### 1.2. function 2: *Wave*
```{r, include=FALSE}
## function 2:
# extract each complete periodic cycle, one node (600x1) at a time
Wave = function(node_data, register){
  # obtain index at which curve crosses 0
  #crossed 0---> -1: pos to neg,    1: neg to pos
  #returns: location index where curve crosses X-axis
  x=diff(ifelse(node_data>0,1,0))       
  z_idx=(1:length(node_data))[x!=0]             
  # skip first crossing if it is from positive to negative
  if (x[z_idx[1]]==-1){
    z_idx=z_idx[-1]
  }
  #put every complete cycle in a Dataframe
  i=1
  cl=1
  result=data.frame(cycle=integer(), time=integer(), y_value=integer())
  while (i+2<=length(z_idx)){
    if(register==0){
      tmp=data.frame(cycle=cl, time=seq(z_idx[i],z_idx[i+2]), y_value=node_data[z_idx[i]:z_idx[i+2]])
    }
    else{
      tmp=data.frame(cycle=cl, time=seq(1,length(seq(z_idx[i],z_idx[i+2]))), y_value=node_data[z_idx[i]:z_idx[i+2]])  
      tmp[,2]=tmp[,2]-length(seq(z_idx[i],z_idx[i+1]))-1
    }
    result=rbind(result,tmp)
    i=i+2
    cl=cl+1
  }
  return(result)
}
```

### 1.3. function 3: *fourier_smooth*
```{r, include=FALSE}
## function 3:
# Defined fourier smoothing functions 
# To study a single brain node response, specify the node number in the *node_subset* list. 
fourier_smooth <- function(data_mat, time_subset, node_subset, k){
  basis <- create.fourier.basis(c(time_subset[1],time_subset[length(time_subset)]), k)
  fd_obj <- smooth.basis(time_subset, data_mat[time_subset,node_subset], basis)
  smoothfd <- fd_obj$fd
  #plot(smoothfd)
  #title(main=paste("Fourier Basis Smoothing of node:", node_subset, ", Basis_number:",k ))
  return(fd_obj)
}
```

### 1.4. function 4: *fPCA.nodes*
```{r, include=FALSE}
## function 4
# FPCA function
fPCA.nodes <- function(data_mat, k, nharm, plt){
  smoothfd <- fourier_smooth(data_mat, time_subset=c(1:nrow(data_mat)),
                             node_subset=c(1:ncol(data_mat)), k)$fd
  pcalist = pca.fd(smoothfd, nharm, harmfdPar=fdPar(smoothfd))
  rotpcalist = varmx.pca.fd(pcalist)
  par(mfrow=c(nharm,1))
  if(plt==1){
    plot.pca.fd(rotpcalist)
  } 
  return(rotpcalist)
}
```

### 1.5. function 5: *node.scaler*
```{r, include=FALSE}
## function 5 
# Scale the node timeframe to [-1,1] and make make sure each cycle has same number of rows
node.scaler <-  function(data_mat, node, k=32){
    result_obj <- fourier_smooth(data_mat, time_subset=c(1:nrow(data_mat)), node_subset=c(node), k)
    smoothed_curve = eval.fd(c(1:nrow(data_mat)),result_obj$fd)
    transformed_node = Wave(smoothed_curve, register=1)
    df_tmp = data.frame(cycle=integer(), time=integer(), y_value=integer())
    for(i in 1:length(unique(transformed_node$cycle))){
      tmp=subset(transformed_node, cycle==i)
      tmp$time = ifelse(tmp$time<=0, (tmp$time)/abs(min(tmp$time)), (tmp$time)/abs(max(tmp$time)))
      df_tmp=rbind(df_tmp,tmp)
    }
    par(mfrow=c(1,1))
    return(df_tmp)
}
```

### 1.6. function 6: *row.check*
```{r, include=FALSE}
## function 6
# make sure they have same row number
row.check <- function(node_data) {
  node_df = data.frame(matrix(nrow=80))
  node.name = unique(node_data$cycle)
  for(i in 1:length(unique(node_data$cycle))){
    xx=seq(-1,1,length.out=80)
    tmp=subset(node_data, cycle==i)
    s=smooth.spline(x=tmp$time, y=tmp$y_value,  df = 10)
    node_df[,ncol(node_df)+1]=predict(s,xx)$y
  }
  node_df=node_df[,2:(length(unique(node_data$cycle))+1)]  
  #node_df <- as.matrix(node_df)
  names(node_df) = node.name
  return(node_df)
}
```

##########################################

##########################################
#         **main  script**              #
##########################################

## 2. Examples of using each function
### 2.1 Read data
```{r}
file_loc1 = '/Users/hanwang/desktop/Git_desktop/Functional_Data_Analysis/data_15may2020.csv'
file_loc5 = '/Users/hanwang/desktop/Git_desktop/Functional_Data_Analysis/Data/Patient5.csv'
file_loc9 = '/Users/hanwang/desktop/Git_desktop/Functional_Data_Analysis/Data/Patient9.csv'
file_loc28 = '/Users/hanwang/desktop/Git_desktop/Functional_Data_Analysis/Data/Patient28.csv'
file_loc80 = '/Users/hanwang/desktop/Git_desktop/Functional_Data_Analysis/Data/Patient80.csv'

patient1 <- ReadFile(file_loc1, time_subset=c(1:600), node_subset=c(1:32))
patient5 <- ReadFile(file_loc5, time_subset=c(1:600), node_subset=c(1:32))
patient9 <- ReadFile(file_loc9, time_subset=c(1:600), node_subset=c(1:32))
patient28 <- ReadFile(file_loc28, time_subset=c(1:600), node_subset=c(1:32))
patient80 <- ReadFile(file_loc80, time_subset=c(1:600), node_subset=c(1:32))
```
read in data from the specified file location

### 2.2 *Wave* function -> single node
```{r}
# extract periodic curves from a single node (before smoothing)
node1 = Wave(patient5[,1], register=1)
ggplot(node1, aes(time, y_value,group=cycle, colour=cycle)) + geom_line() + theme(legend.position="right")+ ggtitle("Original")
```
### 2.3(a) node 1 of Patients 1
```{r}
# apply node.scaler to *single node* and apply row.check match row numbers
a = node.scaler(patient1, node=c(1), k=32)
b = row.check(node_data=a)
b = read.zoo(b, index='index')
autoplot(b, facet = NULL, main='After scaling')
# fPCA
rotpcalist = fPCA.nodes(as.matrix(b), k=11, nharm=2, plt=1)
```

### 2.3(b) node 1 of Patients 9
```{r}
# apply node.scaler to *single node* and apply row.check match row numbers
a = node.scaler(patient9, node=c(1), k=32)
b = row.check(node_data=a)
b = read.zoo(b, index='index')
autoplot(b, facet = NULL, main='After scaling')
# fPCA
rotpcalist = fPCA.nodes(as.matrix(b), k=11, nharm=2, plt=1)
```



## 3. Patient 1, Variance among all nodes on the two principle components
```{r, eval=TRUE}
# fPCA on all nodes
PC1_df = data.frame(matrix(nrow=80))
PC2_df = data.frame(matrix(nrow=80))
mean_df = data.frame(matrix(nrow=80))
for(node in 1:32){
  node.df = node.scaler(patient1, node, k=32)
  node.df = as.matrix(row.check(node.df))
  rotpcalist = fPCA.nodes(data_mat=node.df, k=11, nharm=2, plt=0)
  # PC1 & PC2
  harmfd <- rotpcalist[[1]]
  basisfd <- harmfd$basis
  rangex <- basisfd$rangeval
  x <- seq(rangex[1], rangex[2], length = harmfd$basis$rangeval[2])
  fdmat <- eval.fd(x, harmfd)
  meanmat <- eval.fd(x, rotpcalist$meanfd)
  PC1_df[,ncol(PC1_df)+1]=fdmat[,1]    
  PC2_df[,ncol(PC2_df)+1]=fdmat[,2]
  mean_df[,ncol(mean_df)+1]=meanmat
}
PC1_df = data.frame(PC1_df[,2:(ncol(PC1_df))])
names(PC1_df)=colnames(patient1)
PC2_df = data.frame(PC2_df[,2:(ncol(PC2_df))])
names(PC2_df)=colnames(patient1)
mean_df = data.frame(mean_df[,2:(ncol(mean_df))])
names(mean_df)=colnames(patient1)
## plot
### PC1
z_1 = read.zoo(PC1_df, index='index')
### PC2
z_2 = read.zoo(PC2_df, index='index')
### Mean Functional Curves
z_3 = read.zoo(mean_df, index='index')
autoplot(z_1, facet = NULL, main='PC1')
autoplot(z_2, facet = NULL, main='PC2')
autoplot(z_3, facet = NULL, main='mean curves')
```

## 3(b). Patient 28, Variance among all nodes on the two principle components
```{r, eval=TRUE}
# fPCA on all nodes
PC1_df = data.frame(matrix(nrow=80))
PC2_df = data.frame(matrix(nrow=80))
mean_df = data.frame(matrix(nrow=80))
for(node in 1:32){
  node.df = node.scaler(patient28, node, k=32)
  node.df = as.matrix(row.check(node.df))
  rotpcalist = fPCA.nodes(data_mat=node.df, k=11, nharm=2, plt=0)
  # PC1 & PC2
  harmfd <- rotpcalist[[1]]
  basisfd <- harmfd$basis
  rangex <- basisfd$rangeval
  x <- seq(rangex[1], rangex[2], length = harmfd$basis$rangeval[2])
  fdmat <- eval.fd(x, harmfd)
  meanmat <- eval.fd(x, rotpcalist$meanfd)
  fac1 <- sqrt(rotpcalist$values[1]) 
  fac2 <- sqrt(rotpcalist$values[2]) 
  PC1_df[,ncol(PC1_df)+1]=2*fac1*fdmat[,1]    
  PC2_df[,ncol(PC2_df)+1]=2*fac2*fdmat[,2]
  mean_df[,ncol(mean_df)+1]=meanmat
}
PC1_df = data.frame(PC1_df[,2:(ncol(PC1_df))])
names(PC1_df)=colnames(patient1)
PC2_df = data.frame(PC2_df[,2:(ncol(PC2_df))])
names(PC2_df)=colnames(patient1)
mean_df = data.frame(mean_df[,2:(ncol(mean_df))])
names(mean_df)=colnames(patient1)
## plot
### PC1
z_1 = read.zoo(PC1_df, index='index')
### PC2
z_2 = read.zoo(PC2_df, index='index')
### Mean Functional Curves
z_3 = read.zoo(mean_df, index='index')
autoplot(z_1, facet = NULL, main='PC1')
autoplot(z_2, facet = NULL, main='PC2')
autoplot(z_3, facet = NULL, main='mean curves')
```

