---
title: "HC & PD placebo "
author: "Zach Wang"
date: "9/19/2020"
output:
  html_document: default
  word_document: default
  pdf_document: default
---
# Version Note: Variance on PC1 using "Wave" function to analyze periodic sin wave across patients for selected nodes --> (3,4,8,11,15,29,31)
## 1. Function Implementations
### 1.1 function 1: *ReadFile* and attach packages
```{r setup, include=FALSE}
library(zoo)
library(fda.usc)
library(dplyr)
library(reshape)
library(reshape2)
library(pracma)
library(ggplot2)

## function 1:
# function to read data from specified file location, with respect to Node and Time of interest
ReadFile <- function(file_loc, time_subset, node_subset){
  data_mat <- read.csv(file_loc, header=TRUE)
  #data_mat <- select(data_mat, -c(X)) # input data åhas an extra column "X"
  return(as.matrix(data_mat[time_subset,node_subset]))
}
```

### 1.2. function 2: *Wave*
```{r, include=FALSE}
## function 2:
# extract each complete periodic cycle, one node (600x1) at a time
Wave = function(node_data, register){
  # obtain index at which curve crosses 0
  #crossed 0---> -1: pos to neg,    1: neg to pos
  #returns: location index where curve crosses X-axis
  x=diff(ifelse(node_data>0,1,0))       
  z_idx=(1:length(node_data))[x!=0]             
  # skip first crossing if it is from positive to negative
  if (x[z_idx[1]]==-1){
    z_idx=z_idx[-1]
  }
  #put every complete cycle in a Dataframe
  i=1
  cl=1
  result=data.frame(cycle=integer(), time=integer(), y_value=integer())
  while (i+2<=length(z_idx)){
    if(register==0){
      tmp=data.frame(cycle=cl, time=seq(z_idx[i],z_idx[i+2]), y_value=node_data[z_idx[i]:z_idx[i+2]])
    }
    else{
      tmp=data.frame(cycle=cl, time=seq(1,length(seq(z_idx[i],z_idx[i+2]))), y_value=node_data[z_idx[i]:z_idx[i+2]])  
      tmp[,2]=tmp[,2]-length(seq(z_idx[i],z_idx[i+1]))-1
    }
    # fixed: start & end at 0 issue
    tmp[1,3]=0
    tmp[nrow(tmp)+1,] = c(cl, tmp[nrow(tmp),2]+1, 0)
    result=rbind(result,tmp)
    i=i+2
    cl=cl+1
  }
  return(result)
}
```

### 1.3. function 3: *fourier_smooth*
```{r, include=FALSE}
## function 3:
# Defined fourier smoothing functions 
# To study a single brain node response, specify the node number in the *node_subset* list. 
fourier_smooth <- function(data_mat, time_subset, node_subset, k){
  basis <- create.fourier.basis(c(time_subset[1],time_subset[length(time_subset)]), k)
  fd_obj <- smooth.basis(time_subset, data_mat[time_subset,node_subset], basis)
  smoothfd <- fd_obj$fd
  #plot(smoothfd)
  #title(main=paste("Fourier Basis Smoothing of node:", node_subset, ", Basis_number:",k ))
  return(fd_obj)
}
```

### 1.4. function 4: *fPCA.nodes*
```{r, include=FALSE}
## function 4
# FPCA function
fPCA.nodes <- function(data_mat, k, nharm, plt){
  smoothfd <- fourier_smooth(data_mat, time_subset=c(1:nrow(data_mat)),
                             node_subset=c(1:ncol(data_mat)), k)$fd
  pcalist = pca.fd(smoothfd, nharm, harmfdPar=fdPar(smoothfd))
  rotpcalist = varmx.pca.fd(pcalist)
  par(mfrow=c(nharm,1))
  if(plt==1){
    plot.pca.fd(rotpcalist)
  } 
  return(rotpcalist)
}
```

### 1.5. function 5: *node.scaler*
```{r, include=FALSE}
## function 5 
# Scale the node timeframe to [-1,1] and make make sure each cycle has same number of rows
node.scaler <-  function(data_mat, node, k=32){
    result_obj <- fourier_smooth(data_mat, time_subset=c(1:nrow(data_mat)),
                                 node_subset=c(node), k)
    smoothed_curve = eval.fd(c(1:nrow(data_mat)),result_obj$fd)
    transformed_node = Wave(smoothed_curve, register=1)
    df_tmp = data.frame(cycle=integer(), time=integer(), y_value=integer())
    for(i in 1:length(unique(transformed_node$cycle))){
      tmp=subset(transformed_node, cycle==i)
      tmp$time = ifelse(tmp$time<=0, (tmp$time)/abs(min(tmp$time)), 
                        (tmp$time)/abs(max(tmp$time)))
      df_tmp=rbind(df_tmp,tmp)
    }
    par(mfrow=c(1,1))
    return(df_tmp)
}
```

### 1.6. function 6: *row.check*
```{r, include=FALSE}
## function 6
# make sure they have same row number
row.check <- function(node_data) {
  node_df = data.frame(matrix(nrow=80))
  node.name = unique(node_data$cycle)
  for(i in 1:length(unique(node_data$cycle))){
    xx=seq(-1,1,length.out=80)
    tmp=subset(node_data, cycle==i)
    s=smooth.spline(x=tmp$time, y=tmp$y_value,  df = 10)
    node_df[,ncol(node_df)+1]=predict(s,xx)$y
  }
  node_df=node_df[,2:(length(unique(node_data$cycle))+1)]  
  #node_df <- as.matrix(node_df)
  names(node_df) = node.name
  return(node_df)
}

```

### 1.7 function *pc.df*
it's to put variance on 1st and 2nd Principle Components into dataframe. It is the same.
```{r, include=FALSE}
pc.df <- function(patient_df){
  PC1_df = data.frame(matrix(nrow=80))
  PC2_df = data.frame(matrix(nrow=80))
  mean_df = data.frame(matrix(nrow=80))
  out = list()
  for(node in 1:ncol(patient_df)){
    node.df = node.scaler(patient_df, node, k=21)
    node.df = as.matrix(row.check(node.df))
    rotpcalist = fPCA.nodes(data_mat=node.df, k=11, nharm=2, plt=0)
    # PC1 & PC2
    harmfd <- rotpcalist[[1]]
    basisfd <- harmfd$basis
    rangex <- basisfd$rangeval
    x <- seq(rangex[1], rangex[2], length = harmfd$basis$rangeval[2])
    meanmat <- eval.fd(x, rotpcalist$meanfd)
    fdmat <- eval.fd(x, harmfd)
    fac1 <- sqrt(rotpcalist$values[1]) 
    fac2 <- sqrt(rotpcalist$values[2]) 
    
    PC1_df[,ncol(PC1_df)+1]=2*fac1*fdmat[,1]    
    PC2_df[,ncol(PC2_df)+1]=2*fac2*fdmat[,2]
    mean_df[,ncol(mean_df)+1]=meanmat
  }
  PC1_df = data.frame(PC1_df[,2:(ncol(PC1_df))])
  names(PC1_df)=colnames(patient_df)
  PC2_df = data.frame(PC2_df[,2:(ncol(PC2_df))])
  names(PC2_df)=colnames(patient_df)
  mean_df = data.frame(mean_df[,2:(ncol(mean_df))])
  names(mean_df)=colnames(patient_df)
  out$pc1 = PC1_df
  out$pc2 = PC2_df
  out$mean = mean_df
  return(out)
}
```

##########################################
#         **main  script**              #
##########################################

## 1. Parkinson Disease - Placebo Group
```{r}
pd_placebo <- c(1,4,6,10,14,15,18,20,22,23,31,46,47,50,56,60,61,64,65,80,81,91,97)
for (i in pd_placebo){
  df_tmp = ReadFile(paste('/Users/hanwang/desktop/Git_desktop/Functional_Data_Analysis/Data for Zach/Data for Zach ', i,'.csv', sep=""),
                    time_subset=c(1:600), node_subset=c(1:32))
  tmp=pc.df(df_tmp)
  a =tmp$pc1
  a$index = c(1:80)
  d <- melt(data = a, id.vars = c("index"), measure.vars = colnames(df_tmp))
  print(ggplot(d, aes(x = index, y = value)) + 
          geom_line(aes(color = variable))+ggtitle(paste("PD, patient", i, "PC1")))
}
```

## 2. Healthy Control - Placebo Group
```{r}
hc_placebo <- c(11,25,29,33,38,40,41,43,51,58,67,70,71,73,86,90,94,96,100)
for (i in hc_placebo){
  df_tmp = ReadFile(paste('/Users/hanwang/desktop/Git_desktop/Functional_Data_Analysis/Data for Zach/Data for Zach ', i,'.csv', sep=""),
                    time_subset=c(1:600), node_subset=c(1:32))
  tmp=pc.df(df_tmp)
  a =tmp$pc1
  a$index = c(1:80)
  d <- melt(data = a, id.vars = c("index"), measure.vars = colnames(df_tmp))
  print(ggplot(d, aes(x = index, y = value)) + 
          geom_line(aes(color = variable))+ggtitle(paste("HC, patient", i, "PC1")))
}
```
