---
title: "3aug2020_v1"
author: "Zach Wang"
date: "8/3/2020"
output: html_document
---

## 1. progress summary
This week, as we discussed from last meeting, I have: 
- 1. selected the 1st complete sin curve of every node and align them to register at the origin.
- 2. Scale the the time frame of each node to [-1,1] while time=0 remain unchanged. I also make sure that each node have the same number of rows so that fPCA will be able to apply.
- 3. Apply fPCA to the new dataset of curves (1st cycle of each node), and the result is that the first two Principle Component functions covers 91.5% of total variance. 
- 4. repeat the process above for 2nd cycle of each node, and first two Principle component functions covered in total 93.7% of total variance.

## 2. read data and attach packages
```{r setup, include=FALSE}
library(zoo)
library(fda.usc)
library(dplyr)
library(reshape)
library(reshape2)
library(pracma)
library(ggplot2)

#data_15may2020 <- read.csv("C:/Users/Han Wang/Desktop/BioRankings/data_15may2020.csv", header=TRUE)
data_15may2020 <- read.csv("/Users/hanwang/desktop/Git_desktop/Functional_Data_Analysis/data_15may2020.csv", header=TRUE)
z <- read.zoo(data_15may2020)
data_15may2020 <- select(data_15may2020, -c(X))
data_mat <- as.matrix(data_15may2020)
```

## 3. Defined fourier smoothing functions 
To study a single brain node response, specify the node number in the *node_subset* list. 
```{r smoothing_function}
f_fourier_smooth <- function(time_subset, data_mat, node_subset, k){
  basis <- create.fourier.basis(c(time_subset[1],time_subset[length(time_subset)]), k)
  fd_obj <- smooth.basis(time_subset, data_mat[time_subset,node_subset], basis)
  smoothfd <- fd_obj$fd
  #plot(smoothfd)
  #title(main=paste("Fourier Basis Smoothing of node:", node_subset, ", Basis_number:",k ))
  return(fd_obj)
}
```

## 4. define the function to extract periodic cycle of a single node response
```{r , include=FALSE}
transform.Cycle = function(data, register){
  # obtain index at which curve crosses 0
  #crossed 0---> -1: pos to neg,    1: neg to pos
  #returns: location index where curve crosses X-axis
  x=diff(ifelse(data>0,1,0))       
  z_idx=(1:599)[x!=0]             
  
  # skip first crossing if it is from positive to negative
  if (x[z_idx[1]]==-1){
    z_idx=z_idx[-1]
  }
  
  #put every complete cycle in a Dataframe
  i=1
  cl=1
  result=data.frame(cycle=integer(), time=integer(), y_value=integer())
  while (i+2<=length(z_idx)){
    
      if(register==0){
        tmp=data.frame(cycle=cl, time=seq(z_idx[i],z_idx[i+2]),
                       y_value=smoothed_curve[z_idx[i]:z_idx[i+2]])
      }
      else{
        tmp=data.frame(cycle=cl, time=seq(1,length(seq(z_idx[i],z_idx[i+2]))),
                       y_value=smoothed_curve[z_idx[i]:z_idx[i+2]])  
      }
      result=rbind(result,tmp)
      i=i+2
      cl=cl+1
    }
  return(result)
}

```

## 5. Getting Harmonics curves based on Eigenfunctions
Note: I am having problem getting the math right. 
Below, I calculated each Harmonics based on the formula:
 $$ Eigenfunction(t)=\sum_{k=1}^{K}c_{k} \Phi_{k}(t) $$
 and 
 $$ \Phi_{k}(t)=c_1+c_2sin(t)+c_3cos(t)+c_4sin(t)+c_5cos(t)+...
 =c_1+(c_2+c_4+...)sin(t)+(c_3+c_5+...)cos(t)
 $$
```{r}
# find y value based on PC coef and sin/cos functions
PC1_df = data.frame(matrix(nrow=80))
PC2_df = data.frame(matrix(nrow=80))
for(node in 1:32){
    result_obj <- f_fourier_smooth(time_subset=c(1:600), data_mat, node_subset=c(node), k=32)
    smoothed_curve = eval.fd(c(1:600),result_obj$fd)
    transformed_node = transform.Cycle(smoothed_curve, register=1)
    df_tmp = data.frame(cycle=integer(), time=integer(), y_value=integer())
    for(i in 1:length(unique(transformed_node$cycle))){
      tmp=subset(transformed_node, cycle==i)
      tmp$time=(tmp$time)/max(tmp$time)
      df_tmp=rbind(df_tmp,tmp)
    }
    df_new = data.frame(matrix(nrow=80))
    for(i in 1:length(unique(df_tmp$cycle))){
      xx=seq(0,1,length.out=80)
      tmp=subset(df_tmp, cycle==i)
      s=smooth.spline(x=tmp$time, y=tmp$y_value,  df = 10)
      df_new[,ncol(df_new)+1]=predict(s,xx)$y
    }
    df_new=df_new[,2:(length(unique(df_tmp$cycle))+1)]
    ## FPCA now
    fPCA_subset <- function(data_mat, k, nharm, plt){
      basis <- create.fourier.basis(c(1,nrow(data_mat)), k) 
      smoothfd <- smooth.basis(1:nrow(data_mat), data_mat, basis)$fd 
      pcalist = pca.fd(smoothfd, nharm, harmfdPar=fdPar(smoothfd))
      rotpcalist = varmx.pca.fd(pcalist)
      par(mfrow=c(nharm,1))
      if(plt==1){
        plot.pca.fd(rotpcalist)
      } 
      return(rotpcalist)
    }
    df_new <- as.matrix(df_new)
    rotpcalist = fPCA_subset(df_new, k=11, nharm=2, plt=0)
    PCA_df <- data.frame(matrix(nrow=rotpcalist$harmonics$basis$rangeval[2]))
    # Calculation for Harmonics/eigenvectors
    for(x in (rotpcalist$harmonics$basis$rangeval[1]:rotpcalist$harmonics$basis$rangeval[2])){
      PC1_coef = rotpcalist$harmonics$coefs[,1]
      PC2_coef = rotpcalist$harmonics$coefs[,2]
      flag = ifelse(index(PC1_coef)%%2==0, 1, 0)
      even_idx = (1:length(PC1_coef))[flag==1]
      odd_idx = (1:length(PC1_coef))[flag==0][-1]
      PCA_df[x, 1] = PC1_coef[1] + sin(x)*sum(PC1_coef[even_idx]) + cos(x)*sum(PC1_coef[odd_idx]) # fixme
      PCA_df[x, 2] = PC2_coef[1] + sin(x)*sum(PC2_coef[even_idx]) + cos(x)*sum(PC2_coef[odd_idx]) # fixme
    
        ## calculation trial 2
      #PCA_df[x, 1] = PC1_coef[1]
      #PCA_df[x, 2] = PC2_coef[1]
      #t=1
      #for(i in 1:length((even_idx))){
      #  PCA_df[x, 1]=PCA_df[x, 1]+sin(t*x*PC1_coef[even_idx[i]])
      #  PCA_df[x, 2]=PCA_df[x, 2]+sin(t*x*PC2_coef[even_idx[i]])
      #  t=t+1
      #}
      #t=1
      #for(j in 1:length((odd_idx))){
      #  PCA_df[x, 1]=PCA_df[x, 1]+sin(t*x*PC1_coef[odd_idx[j]])
      #  PCA_df[x, 2]=PCA_df[x, 2]+sin(t*x*PC2_coef[odd_idx[j]])
      #  t=t+1
      #}
      
    }
PC1_df[,ncol(PC1_df)+1]=PCA_df[,1]    
PC2_df[,ncol(PC2_df)+1]=PCA_df[,2]
}

PC1_df = data.frame(PC1_df[,2:(ncol(PC1_df))])
names(PC1_df)=colnames(data_mat)

PC2_df = data.frame(PC2_df[,2:(ncol(PC2_df))])
names(PC2_df)=colnames(data_mat)

z_1 = read.zoo(PC1_df, index='index')
z_2 = read.zoo(PC2_df, index='index')
autoplot(z_1, facet = NULL)
```

## 6. Plot the 32 Harmonics plots directly instead 
```{r}
# plot 32 plots instead
for(node in 1:32){
    result_obj <- f_fourier_smooth(time_subset=c(1:600), data_mat, node_subset=c(node), k=32)
    smoothed_curve = eval.fd(c(1:600),result_obj$fd)
    transformed_node = transform.Cycle(smoothed_curve, register=1)
    df_tmp = data.frame(cycle=integer(), time=integer(), y_value=integer())
    for(i in 1:length(unique(transformed_node$cycle))){
      tmp=subset(transformed_node, cycle==i)
      tmp$time=(tmp$time)/max(tmp$time)
      df_tmp=rbind(df_tmp,tmp)
    }
    df_new = data.frame(matrix(nrow=80))
    for(i in 1:length(unique(df_tmp$cycle))){
      xx=seq(0,1,length.out=80)
      tmp=subset(df_tmp, cycle==i)
      s=smooth.spline(x=tmp$time, y=tmp$y_value,  df = 10)
      df_new[,ncol(df_new)+1]=predict(s,xx)$y
    }
    df_new=df_new[,2:(length(unique(df_tmp$cycle))+1)]
    ## FPCA now
    fPCA_subset <- function(data_mat, k, nharm, plt){
      basis <- create.fourier.basis(c(1,nrow(data_mat)), k) 
      smoothfd <- smooth.basis(1:nrow(data_mat), data_mat, basis)$fd 
      pcalist = pca.fd(smoothfd, nharm, harmfdPar=fdPar(smoothfd))
      rotpcalist = varmx.pca.fd(pcalist)
      par(mfrow=c(nharm,1))
      if(plt==1){
        plot.pca.fd(rotpcalist)
      } 
      return(rotpcalist)
    }
    df_new <- as.matrix(df_new)
    rotpcalist = fPCA_subset(df_new, k=11, nharm=2, plt=0)
    par(mfrow=c(1,1))
    plot(rotpcalist$harmonics)
    title(paste("node",node))
}
```