---
title: "Function documentations & Examples"
author: "Zach Wang"
date: "8/19/2020"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
## 1. Function Implementations
### 1.1 function 1: *ReadFile* and attach packages
```{r setup, include=FALSE}
library(zoo)
library(fda.usc)
library(dplyr)
library(reshape)
library(reshape2)
library(pracma)
library(ggplot2)

## function 1:
# function to read data from specified file location, with respect to Node and Time of interest
ReadFile <- function(file_loc, time_subset, node_subset){
  data_mat <- read.csv(file_loc, header=TRUE)
  data_mat <- select(data_mat, -c(X)) # input data åhas an extra column "X"
  return(as.matrix(data_mat[time_subset,node_subset]))
}
```

### 1.2. function 2: *Wave*
```{r, include=FALSE}
## function 2:
# extract each complete periodic cycle, one node (600x1) at a time
Wave = function(node_data, register){
  # obtain index at which curve crosses 0
  #crossed 0---> -1: pos to neg,    1: neg to pos
  #returns: location index where curve crosses X-axis
  x=diff(ifelse(node_data>0,1,0))       
  z_idx=(1:599)[x!=0]             
  # skip first crossing if it is from positive to negative
  if (x[z_idx[1]]==-1){
    z_idx=z_idx[-1]
  }
  #put every complete cycle in a Dataframe
  i=1
  cl=1
  result=data.frame(cycle=integer(), time=integer(), y_value=integer())
  while (i+2<=length(z_idx)){
    if(register==0){
      tmp=data.frame(cycle=cl, time=seq(z_idx[i],z_idx[i+2]), y_value=node_data[z_idx[i]:z_idx[i+2]])
    }
    else{
      tmp=data.frame(cycle=cl, time=seq(1,length(seq(z_idx[i],z_idx[i+2]))), y_value=node_data[z_idx[i]:z_idx[i+2]])  
      tmp[,2]=tmp[,2]-length(seq(z_idx[i],z_idx[i+1]))-1
    }
    result=rbind(result,tmp)
    i=i+2
    cl=cl+1
  }
  return(result)
}
```

### 1.3. function 3: *fourier_smooth*
```{r, include=FALSE}
## function 3:
# Defined fourier smoothing functions 
# To study a single brain node response, specify the node number in the *node_subset* list. 
fourier_smooth <- function(data_mat, time_subset, node_subset, k){
  basis <- create.fourier.basis(c(time_subset[1],time_subset[length(time_subset)]), k)
  fd_obj <- smooth.basis(time_subset, data_mat[time_subset,node_subset], basis)
  smoothfd <- fd_obj$fd
  #plot(smoothfd)
  #title(main=paste("Fourier Basis Smoothing of node:", node_subset, ", Basis_number:",k ))
  return(fd_obj)
}
```

### 1.4. function 4: *fPCA.nodes*
```{r, include=FALSE}
## function 4
# FPCA function
fPCA.nodes <- function(data_mat, k, nharm, plt){
  smoothfd <- fourier_smooth(data_mat, time_subset=c(1:nrow(data_mat)), node_subset=c(1:ncol(data_mat)), k)$fd
  pcalist = pca.fd(smoothfd, nharm, harmfdPar=fdPar(smoothfd))
  rotpcalist = varmx.pca.fd(pcalist)
  par(mfrow=c(nharm,1))
  if(plt==1){
    plot.pca.fd(rotpcalist)
  } 
  return(rotpcalist)
}
```

### 1.5. function 5: *node.scaler*
```{r, include=FALSE}
## function 5 
# Scale the node timeframe to [-1,1] and make make sure each cycle has same number of rows
node.scaler <-  function(data_mat, node){
    result_obj <- fourier_smooth(data_mat, time_subset=c(1:600), node_subset=c(node), k=32)
    smoothed_curve = eval.fd(c(1:600),result_obj$fd)
    transformed_node = Wave(smoothed_curve, register=1)
    df_tmp = data.frame(cycle=integer(), time=integer(), y_value=integer())
    for(i in 1:length(unique(transformed_node$cycle))){
      tmp=subset(transformed_node, cycle==i)
      tmp$time = ifelse(tmp$time<=0, (tmp$time)/abs(min(tmp$time)), (tmp$time)/abs(max(tmp$time)))
      df_tmp=rbind(df_tmp,tmp)
    }
    par(mfrow=c(1,1))
    return(df_tmp)
}
```

### 1.6. function 6: *row.check*
```{r, include=FALSE}
## function 6
# make sure they have same row number
row.check <- function(node_data) {
  node_df = data.frame(matrix(nrow=80))
  node.name = unique(node_data$cycle)
  for(i in 1:length(unique(node_data$cycle))){
    xx=seq(-1,1,length.out=80)
    tmp=subset(node_data, cycle==i)
    s=smooth.spline(x=tmp$time, y=tmp$y_value,  df = 10)
    node_df[,ncol(node_df)+1]=predict(s,xx)$y
  }
  node_df=node_df[,2:(length(unique(node_data$cycle))+1)]  
  #node_df <- as.matrix(node_df)
  names(node_df) = node.name
  return(node_df)
}
```

##########################################
#         **main  script**              #
##########################################

## 2. Examples of using each function
### 2.1 Read data
```{r}
file_loc = '/Users/hanwang/desktop/Git_desktop/Functional_Data_Analysis/data_15may2020.csv'
data_15may2020 <- ReadFile(file_loc, time_subset=c(1:600), node_subset=c(1:32))
```
read in data from the specified file location

### 2.2 *Wave* function
**Wave** function takes the a single node data as input, and transforms the data to N complete sinusoidal cycles.
Below are some examples of using Wave function on data of a single node.

* **Notes:**
  + *Wave* has two arguements:
    - node_data: input takes one column at a time (here is a 600x1 matrix)
    - register: set to 0 if want the result to be the original curves; set to 1 if want the result curves to center around 0 

The following 2 plots are generated by the *Wave* function applied on the raw data of node 1 (before smoothing).
```{r}
# extract periodic curves from a single node (before smoothing)
node1 = Wave(data_15may2020[,1], register=0)
ggplot(node1, aes(time, y_value,group=cycle, colour=cycle)) + geom_line() + theme(legend.position="top")
node1 = Wave(data_15may2020[,1], register=1)
ggplot(node1, aes(time, y_value,group=cycle, colour=cycle)) + geom_line() + theme(legend.position="top")
```

The following 2 plots are generated by *Wave* functions on the transformed data of node 1 (after smoothing).
In the examples below, another function, **fourier_smooth** is called first to smooth the raw data using fourier basis functions.

* **Notes:**
  + *fourier_smooth* has four arguments:
    - data_mat: Input data matrix, which should contains the entire data set, but we can use other arguments to select only *"subsets"* or *"Region of interest"* of the data for analysis
    - time_subset: specify a subset of time or all for analysis
    - node_subset: specify a subset of node or all for analysis
    - k: specify the *"nummber of fourier basis functions"* to be used in the smoothing process
```{r}
# extract periodic curves from a single node (after smoothing)
result_obj <- fourier_smooth(data_mat=data_15may2020, time_subset=c(1:600), node_subset=c(1), k=32)
node1_smoothed = eval.fd(c(1:600),result_obj$fd)
node1_smoothed_extraced = Wave(node1_smoothed, register=0)
ggplot(node1_smoothed_extraced, aes(time, y_value,group=cycle, colour=cycle)) + geom_line() + theme(legend.position="top")
node1_smoothed_extraced = Wave(node1_smoothed, register=1)
ggplot(node1_smoothed_extraced, aes(time, y_value,group=cycle, colour=cycle)) + geom_line() + theme(legend.position="top")
```

### 2.3 *fPCA.nodes* function
**fPCA.nodes** function takes input data in matrix form (N rows x M columns), and transform the data to N complete sinusoidal cycles -- there has to be more than one column, otherwise PCA won't make sense.

* **Notes:**
  + *fPCA.nodes* has four arguements:
    - data_mat: data in matrix form (N rows x M columns).
    - k: number of fourier basis functions used
    - nharm: number of Principle Components interested
    - plt: set to 1 if want to generate plot; set to 0 if don't want plot
    
Below are some examples of using **fPCA.nodes** function on the original data.
```{r}
# fPCA on raw data
rotpcalist = fPCA.nodes(data_15may2020, k=11, nharm=2, plt=1)
```

### 2.4 *node.scaler* & *row.check* functions
**node.scaler** function is designed to transform a single node curves into T complete sinusoidal cycles (*Wave* function is called). The returned matrix has 3 columns (*cycle*, *time*, *y_value*). The *time* columns will be scaled between range of [-1,1]. 

* **Notes:**
  + *node.scaler* has two arguements: 
    - data_mat: input data in matrix form (N rows x M columns).
    - node: specify which node will be currently used, since scaler function only look at one node at a time
    
**row.check** function is designed to transform the returned dataframe from calling the *node.scaler* function, and most importantly make sure the total number of rows is the same for every column. The result is a 80 x T matrix (the pivot of the matrix from *node.scaler* result), so that each column now represents a different *cycle* and each row stores the *y_value* corresponding to each time point.

* **Notes:**
  + *row.check*  has only one input argument:
    - node_data: the transformed node data from *node.scaler*.
    
Below are examples of using **node.scaler** together with **row.check** function on a single node.
Two plots were generated to show that the result shape of curves after calling the node.scaler and row.check functions are the same.
```{r}
# apply scaler to *single node* and match row numbers
a = node.scaler(data_15may2020, node=c(1))
ggplot(a, aes(time, y_value,group=cycle, colour=cycle)) + geom_line() + theme(legend.position="top")
b = row.check(node_data=a)
b = read.zoo(b, index='index')
autoplot(b, facet = NULL)
```


## 3. Use all the functions defined above to analyze the entire dataset (all nodes, all time points)

```{r, eval=TRUE}
# fPCA on all nodes
PC1_df = data.frame(matrix(nrow=80))
PC2_df = data.frame(matrix(nrow=80))
mean_df = data.frame(matrix(nrow=80))
for(node in 1:32){
  node.df = node.scaler(data_15may2020, node)
  node.df = as.matrix(row.check(node.df))
  rotpcalist = fPCA.nodes(data_mat=node.df, k=11, nharm=2, plt=0)
  # PC1 & PC2
  harmfd <- rotpcalist[[1]]
  basisfd <- harmfd$basis
  rangex <- basisfd$rangeval
  x <- seq(rangex[1], rangex[2], length = harmfd$basis$rangeval[2])
  fdmat <- eval.fd(x, harmfd)
  meanmat <- eval.fd(x, rotpcalist$meanfd)
  PC1_df[,ncol(PC1_df)+1]=fdmat[,1]    
  PC2_df[,ncol(PC2_df)+1]=fdmat[,2]
  mean_df[,ncol(mean_df)+1]=meanmat
}
PC1_df = data.frame(PC1_df[,2:(ncol(PC1_df))])
names(PC1_df)=colnames(data_15may2020)
PC2_df = data.frame(PC2_df[,2:(ncol(PC2_df))])
names(PC2_df)=colnames(data_15may2020)
mean_df = data.frame(mean_df[,2:(ncol(mean_df))])
names(mean_df)=colnames(data_15may2020)
## plot
### PC1
z_1 = read.zoo(PC1_df, index='index')
### PC2
z_2 = read.zoo(PC2_df, index='index')
### Mean Functional Curves
z_3 = read.zoo(mean_df, index='index')
autoplot(z_1, facet = NULL)
autoplot(z_2, facet = NULL)
autoplot(z_3, facet = NULL)
```

