---
title: "24july2020_p2"
author: "Zach Wang"
date: "7/24/2020"
output: html_document
---

## 1. progress summary
This week, as we discussed from last meeting, I have: 
- 1. selected the 1st complete sin curve of every node and align them to register at the origin.
- 2. Scale the the time frame of each node to [-1,1] while time=0 remain unchanged. I also make sure that each node have the same number of rows so that fPCA will be able to apply.
- 3. Apply fPCA to the new dataset of curves (1st cycle of each node), and the result is that the first two Principle Component functions covers 91.5% of total variance. 
- 4. repeat the process above for 2nd cycle of each node, and first two Principle component functions covered in total 93.7% of total variance.

## 2. read data and attach packages
```{r setup, include=FALSE}
library(zoo)
library(fda.usc)
library(dplyr)
library(reshape)
library(reshape2)
library(pracma)
library(ggplot2)

#data_15may2020 <- read.csv("C:/Users/Han Wang/Desktop/BioRankings/data_15may2020.csv", header=TRUE)
data_15may2020 <- read.csv("/Users/hanwang/desktop/Git_desktop/Functional_Data_Analysis/data_15may2020.csv", header=TRUE)
z <- read.zoo(data_15may2020)
data_15may2020 <- select(data_15may2020, -c(X))
data_mat <- as.matrix(data_15may2020)
```

## 3. Defined fourier smoothing functions 
To study a single brain node response, specify the node number in the *node_subset* list. 
```{r smoothing_function}
f_fourier_smooth <- function(time_subset, data_mat, node_subset, k){
  basis <- create.fourier.basis(c(time_subset[1],time_subset[length(time_subset)]), k)
  fd_obj <- smooth.basis(time_subset, data_mat[time_subset,node_subset], basis)
  smoothfd <- fd_obj$fd
  #plot(smoothfd)
  #title(main=paste("Fourier Basis Smoothing of node:", node_subset, ", Basis_number:",k ))
  return(fd_obj)
}
```

## 4. define the function to extract periodic cycle of a single node response
```{r , include=FALSE}
transform.Cycle = function(data, register){
  # obtain index at which curve crosses 0
  #crossed 0---> -1: pos to neg,    1: neg to pos
  #returns: location index where curve crosses X-axis
  x=diff(ifelse(data>0,1,0))       
  z_idx=(1:599)[x!=0]             
  
  # skip first crossing if it is from positive to negative
  if (x[z_idx[1]]==-1){
    z_idx=z_idx[-1]
  }
  
  #put every complete cycle in a Dataframe
  i=1
  cl=1
  result=data.frame(cycle=integer(), time=integer(), y_value=integer())
  while (i+2<=length(z_idx)){
    
      if(register==0){
        tmp=data.frame(cycle=cl, time=seq(z_idx[i],z_idx[i+2]),
                       y_value=smoothed_curve[z_idx[i]:z_idx[i+2]])
      }
      else{
        tmp=data.frame(cycle=cl, time=seq(1,length(seq(z_idx[i],z_idx[i+2]))),
                       y_value=smoothed_curve[z_idx[i]:z_idx[i+2]])  
      }
      result=rbind(result,tmp)
      i=i+2
      cl=cl+1
    }
  return(result)
}

```

## 5.1 single node (node=1), of the entire time frame
```{r 1st_node, include=FALSE}
## single node (node=1), of the entire time frame
result_obj <- f_fourier_smooth(time_subset=c(1:600), data_mat, node_subset=c(1), k=32)
smoothed_curve = eval.fd(c(1:600),result_obj$fd)
transformed_node = transform.Cycle(smoothed_curve, register=1)
ggplot(transformed_node, aes(time, y_value,group=cycle, colour=cycle)) + geom_line() + theme(legend.position="top")

```



##### 5.2 Scale the timeframe of each node
```{r scale_1st_node, include=FALSE}
df_tmp = data.frame(cycle=integer(), time=integer(), y_value=integer())
for(i in 1:length(unique(transformed_node$cycle))){
  tmp=subset(transformed_node, cycle==i)
  tmp$time=(tmp$time)/max(tmp$time)
  df_tmp=rbind(df_tmp,tmp)
}
ggplot(df_tmp, aes(time, y_value,group=cycle, colour=cycle)) + geom_line() + theme(legend.position="top")
```

##### 5.3 pivot the new dataframe
```{r pivot_df, include=FALSE}
df_new = data.frame(matrix(nrow=40))
for(i in 1:length(unique(df_tmp$cycle))){
  xx=seq(0,1,length.out=40)
  tmp=subset(df_tmp, cycle==i)
  s=smooth.spline(x=tmp$time, y=tmp$y_value,  df = 10)
  df_new[,ncol(df_new)+1]=predict(s,xx)$y
}
df_new=df_new[,2:(length(unique(df_tmp$cycle))+1)]
```

##### 5.4 fPCA on new dataframe
```{r}
fPCA_subset <- function(data_mat, k, nharm){
  basis <- create.fourier.basis(c(1,nrow(data_mat)), k) 
  smoothfd <- smooth.basis(1:nrow(data_mat), data_mat, basis)$fd 
  plot(smoothfd)
  title(main="smoothed curves")
  pcalist = pca.fd(smoothfd, nharm, harmfdPar=fdPar(smoothfd))
  rotpcalist = varmx.pca.fd(pcalist)
  par(mfrow=c(nharm,1))
  plot.pca.fd(rotpcalist)
  return(rotpcalist)
}
df_new <- as.matrix(df_new)
rotpcalist = fPCA_subset(df_new, k=11, nharm=2)
```


## 6.1 single node (node=2), of the entire time frame
```{r 2nd_node, include=FALSE}
## single node (node=1), of the entire time frame
result_obj <- f_fourier_smooth(time_subset=c(1:600), data_mat, node_subset=c(2), k=32)
smoothed_curve = eval.fd(c(1:600),result_obj$fd)
transformed_node = transform.Cycle(smoothed_curve, register=1)
ggplot(transformed_node, aes(time, y_value,group=cycle, colour=cycle)) + geom_line() + theme(legend.position="top")

```



##### 6.2 Scale the timeframe of each node
```{r scale_2nd_node, include=FALSE}
df_tmp = data.frame(cycle=integer(), time=integer(), y_value=integer())
for(i in 1:length(unique(transformed_node$cycle))){
  tmp=subset(transformed_node, cycle==i)
  tmp$time=(tmp$time)/max(tmp$time)
  df_tmp=rbind(df_tmp,tmp)
}
ggplot(df_tmp, aes(time, y_value,group=cycle, colour=cycle)) + geom_line() + theme(legend.position="top")
```

##### 6.3 pivot the new dataframe
```{r pivot_df_2, include=FALSE}
df_new = data.frame(matrix(nrow=40))
for(i in 1:length(unique(df_tmp$cycle))){
  xx=seq(0,1,length.out=40)
  tmp=subset(df_tmp, cycle==i)
  s=smooth.spline(x=tmp$time, y=tmp$y_value,  df = 10)
  df_new[,ncol(df_new)+1]=predict(s,xx)$y
}
df_new=df_new[,2:(length(unique(df_tmp$cycle))+1)]
```

##### 6.4 fPCA on new dataframe
```{r}
fPCA_subset <- function(data_mat, k, nharm){
  basis <- create.fourier.basis(c(1,nrow(data_mat)), k) 
  smoothfd <- smooth.basis(1:nrow(data_mat), data_mat, basis)$fd 
  plot(smoothfd)
  title(main="smoothed curves")
  pcalist = pca.fd(smoothfd, nharm, harmfdPar=fdPar(smoothfd))
  rotpcalist = varmx.pca.fd(pcalist)
  par(mfrow=c(nharm,1))
  plot.pca.fd(rotpcalist)
  return(rotpcalist)
}
df_new <- as.matrix(df_new)
rotpcalist = fPCA_subset(df_new, k=11, nharm=2)
```


## 7.1 single node (node=3), of the entire time frame
```{r 3rd_node, include=FALSE}
## single node (node=1), of the entire time frame
result_obj <- f_fourier_smooth(time_subset=c(1:600), data_mat, node_subset=c(3), k=32)
smoothed_curve = eval.fd(c(1:600),result_obj$fd)
transformed_node = transform.Cycle(smoothed_curve, register=1)
ggplot(transformed_node, aes(time, y_value,group=cycle, colour=cycle)) + geom_line() + theme(legend.position="top")

```



##### 7.2 Scale the timeframe of each node
```{r scale_3rd_node, include=FALSE}
df_tmp = data.frame(cycle=integer(), time=integer(), y_value=integer())
for(i in 1:length(unique(transformed_node$cycle))){
  tmp=subset(transformed_node, cycle==i)
  tmp$time=(tmp$time)/max(tmp$time)
  df_tmp=rbind(df_tmp,tmp)
}
ggplot(df_tmp, aes(time, y_value,group=cycle, colour=cycle)) + geom_line() + theme(legend.position="top")
```

##### 7.3 pivot the new dataframe
```{r pivot_df_3, include=FALSE}
df_new = data.frame(matrix(nrow=40))
for(i in 1:length(unique(df_tmp$cycle))){
  xx=seq(0,1,length.out=40)
  tmp=subset(df_tmp, cycle==i)
  s=smooth.spline(x=tmp$time, y=tmp$y_value,  df = 10)
  df_new[,ncol(df_new)+1]=predict(s,xx)$y
}
df_new=df_new[,2:(length(unique(df_tmp$cycle))+1)]
```

##### 7.4 fPCA on new dataframe
```{r}
fPCA_subset <- function(data_mat, k, nharm){
  basis <- create.fourier.basis(c(1,nrow(data_mat)), k) 
  smoothfd <- smooth.basis(1:nrow(data_mat), data_mat, basis)$fd 
  plot(smoothfd)
  title(main="smoothed curves")
  pcalist = pca.fd(smoothfd, nharm, harmfdPar=fdPar(smoothfd))
  rotpcalist = varmx.pca.fd(pcalist)
  par(mfrow=c(nharm,1))
  plot.pca.fd(rotpcalist)
  return(rotpcalist)
}
df_new <- as.matrix(df_new)
rotpcalist = fPCA_subset(df_new, k=11, nharm=2)
```
